---
title: "Project 2 Presentation"
author: "Hannah Dean"
date: "2024-11-11"
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
subtitle: Reds Transactions
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```


## Intro
 - I am analyzing the Cincinnati Reds transactions over the past 10 yrs.
 - Resources: <https://www.mlb.com/reds/roster/transactions>
 - Data: **Reds Transactions**
```{r, echo = F, warning=FALSE, message = FALSE}
library(tseries)
library(vars)
library(forecast)
library(readr)

# Importing my data
mydata <- read_csv("Reds10.csv")

mydata$`# of transactions` = replace(mydata$`# of transactions`, c(53,54), 0)
time_series_data <- ts(mydata$`# of transactions`, frequency = 12, start = c(2014, 1))


```

## Time Series Characteristics

- My data is tracking the number of transactions the Reds make each month on their 40-man roster. 
    - signing
    - activating
    - trading
    - placing on IL
    - optioning
    - etc.
- There are a couple variables that could play a role in impacting the results of how many transactions there are each month:
    - the trade deadlines
    - the pre-season transactions
    - season injuries
    - etc.


## Decomposition Plot
```{r, out.width="50%", fig.align="center"}
decomposed_data = decompose(time_series_data)
plot(decomposed_data)
```

- Trend: There is slight decreasing trend in the data
- Seasonality: There is also yearly seasonality in the data

## Autocorrelation Function
```{r acf, out.width="50%", fig.align="center"}
# sample autocorrelation function
acf <- Acf(time_series_data)
```

This shows us that there is significant yearly seasonality in the data due to the autocorrelation surpassing the bands.

## Decomposition Plot Continued
```{r adf}
adf.test(time_series_data)
```

- My data set is stationary do to the constant mean value over time and little variance increase in the data
- We can also see that according to the Augmented Dickey-Fuller Test my data rejects the Null Hypothesis with a p-value of .01, therefore concluding my data is stationary

Because there is some trend and seasonality and because my data is stationary I decided to use the Holt-Winters Model


## Forecasting Model Choice (Holt Winter Model)

- I chose Exponential Smoothing because my data is stationary

- Out of all three exponential smoothing models I chose the Holt Winter Model because there is both seasonality and slight trend


```{r, echo = FALSE}
library(forecast) 
hw_model = hw(time_series_data, h = 12)
```

$$
\begin{align*}
  \hat{y}_{t+h|t} &= (\ell_{t} + hb_{t})s_{t+h-m(k+1)} \\
  \ell_{t} &= .0849 \frac{y_{t}}{s_{t-m}} + (1 - .0849)(\ell_{t-1} + b_{t-1})\\
  b_{t} &= .0001^*(\ell_{t}-\ell_{t-1}) + (1 - .0001^*)b_{t-1}                \\
  s_{t} &= .0001 \frac{y_{t}}{(\ell_{t-1} + b_{t-1})} + (1 - .0001)s_{t-m}
\end{align*}
$$

## Slide 4: Model Results and Forecast Interpretation 

```{r, out.width="50%", fig.align="center"}
plot(forecast(hw_model), ylim = c(0,100))
forecast(hw_model, h = 12)
```

- My forecasted values show the same general trend in seasonality as my historical data however I would expect some of the numbers to have larger numbers than what was predicted.



## Slide 5: Model Diagnostics and Accuracy Evaluation
```{r accuracy}
accuracy(forecast(hw_model))
ses_model <- ses(time_series_data, h= 12)
accuracy(forecast(ses_model))
holt_model <- holt(time_series_data, h = 12)
accuracy(forecast(holt_model, h = 12))
```

- RMSE: Takes into consideration the large error variables in the prediction process therefore predictions could be off by 16 transactions a month
    - For Simple Exponential Smoothing and Holts Linear Model the predictions are off by 23 transactions a month
- MAE: Predictions are off by 12 transactions a month
    - For Simple Exponential Smoothing and Holts Linear Model the predictions are off by 20 transactions a month
- MASE: This model is more accurate than a naive model
    - For Simple Exponential Smoothing and Holts Linear Model the value for MASE surpasses 1 meaning that naive model would be more accurate
- ACF1: Little correlation in errors
    - For Simple Exponential Smoothing and Holts Linear Model correlation in errors is relitevely the same as the Holt Winters



