---
title: "Cross Validation"
author: "Hannah Dean"
date: "2026-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)

CDI_outcomes_2025_2026 <- read_csv(
  "CDI_Full_Data_With_2025_2026_Outcomes.csv",
  show_col_types = FALSE
)
```

## 10-fold cross-validation

```{r cross val}
library(caret)
library(dplyr)
library(purrr)
library(tibble)

set.seed(123)

#----------------------------------------
# Outcomes
#----------------------------------------

outcomes <- c(
  "pb_xRV100_2025",
  "observed_RV100_2025",
  "ERA_2025",
  "SIERA_2025",
  "xERA_2025",
  "FIP_2025",
  "xFIP_2025",
  "RA9_Wins_2025",
  "pb_xRV100_2026",
  "SIERA_2026",
  "xERA_2026",
  "FIP_2026",
  "xFIP_2026",
  "ERA_2026"
)

#----------------------------------------
# Cross-validation settings
#----------------------------------------

ctrl <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 25
)

#----------------------------------------
# Function
#----------------------------------------

cross_validate_model <- function(outcome){

  model_data <-
    CDI_outcomes_2025_2026 %>%
    select(
      all_of(outcome),
      z_stuff,
      z_command,
      z_miss,
      z_contact
    ) %>%
    na.omit()

  fit <- train(
    reformulate(
      c(
        "z_stuff",
        "z_command",
        "z_miss",
        "z_contact"
      ),
      response = outcome
    ),
    data = model_data,
    method = "lm",
    trControl = ctrl,
    metric = "RMSE"
  )

  tibble(
    Outcome = outcome,
    N = nrow(model_data),
    RMSE = fit$results$RMSE,
    Rsquared = fit$results$Rsquared,
    MAE = fit$results$MAE
  )

}

#----------------------------------------
# Run every model
#----------------------------------------

cv_results <-
  map_dfr(outcomes, cross_validate_model)

cv_results
```

Cross-validation demonstrated that the CDI framework generalized well across multiple run-prevention metrics. The highest predictive performance was observed for the 2025 skill-based metrics, including xERA (R² = 0.715), SIERA (R² = 0.693), xFIP (R² = 0.567), FIP (R² = 0.508), and PitchBot xRV100 (R² = 0.533), indicating that the model maintained strong predictive accuracy on unseen data. Prediction accuracy was lower for traditional outcomes such as ERA (R² = 0.383) and RA9-Wins (R² = 0.162), which are influenced by contextual factors including defense, sequencing, and run support.
