---
title: "Plus metrics validation"
author: "Hannah Dean"
date: "2026-07-11"
output:
  word_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(ggplot2)
library(broom)
library(car)
library(rsample)
library(yardstick)
library(MASS)
library(corrplot)
library(knitr)

set.seed(42)
```

```{r load-data}
P_stats_with_stuff <- readRDS(
  "data/P_stats_with_stuff.rds"
)

Command_Metrics_Table <- readRDS(
  "data/Command_Metrics_Table.rds"
)

Miss_Ability_Table <- readRDS(
  "data/Miss_Ability_Table.rds"
)

Contact_Metrics_Table <- readRDS(
  "data/Contact_Metrics_Table.rds"
)

PBP <- readRDS(
  "data/PBP.rds"
)

P_stats <- readRDS(
  "data/P_stats.rds"
)
```

```{r master analysis}
# ============================================================
# CREATE MASTER ELITE TEST TABLE
# ============================================================

elite_test <- P_stats_with_stuff %>%

  # Fangraphs / pitching model metrics
  dplyr::select(
    pitcher,
    player_full_name,
    stuff_plus,
    location_plus,
    pitching_plus
  ) %>%

  dplyr::distinct(
    pitcher,
    .keep_all = TRUE
  ) %>%

  # Add command score
  dplyr::left_join(
    Command_Metrics_Table %>%
      dplyr::select(
        pitcher,
        command_score
      ) %>%
      dplyr::distinct(
        pitcher,
        .keep_all = TRUE
      ),
    by = "pitcher"
  ) %>%

  # Add miss ability score
  dplyr::left_join(
    Miss_Ability_Table %>%
      dplyr::select(
        pitcher,
        miss_score
      ) %>%
      dplyr::distinct(
        pitcher,
        .keep_all = TRUE
      ),
    by = "pitcher"
  ) %>%

  # Add contact management score
  dplyr::left_join(
    Contact_Metrics_Table %>%
      dplyr::select(
        pitcher,
        contact_score
      ) %>%
      dplyr::distinct(
        pitcher,
        .keep_all = TRUE
      ),
    by = "pitcher"
  ) %>%

  # ==========================================================
  # STANDARDIZE ALL COMPONENTS
  # ==========================================================

  dplyr::mutate(
    dplyr::across(
      c(
        stuff_plus,
        location_plus,
        pitching_plus,
        command_score,
        miss_score,
        contact_score
      ),
      ~ as.numeric(scale(.x)),
      .names = "z_{.col}"
    )
  ) %>%

  # Rename standardized columns
  dplyr::rename(
    z_stuff = z_stuff_plus,
    z_location = z_location_plus,
    z_pitching = z_pitching_plus,
    z_command = z_command_score,
    z_miss = z_miss_score,
    z_contact = z_contact_score
  ) %>%

  # ==========================================================
  # CREATE ELITE SCORE VERSIONS
  # ==========================================================

  dplyr::mutate(

    # Original model using Stuff+
    elite_original =
      0.30 * z_stuff +
      0.25 * z_command +
      0.25 * z_miss +
      0.20 * z_contact,

    # Replace Stuff+ with Pitching+
    elite_pitching_only =
      0.30 * z_pitching +
      0.25 * z_command +
      0.25 * z_miss +
      0.20 * z_contact,

    # Split the original 30% between Stuff+ and Location+
    elite_stuff_location =
      0.18 * z_stuff +
      0.12 * z_location +
      0.25 * z_command +
      0.25 * z_miss +
      0.20 * z_contact,

    # Split the original 30% among Stuff+, Location+, and Pitching+
    elite_all_three =
      0.12 * z_stuff +
      0.08 * z_location +
      0.10 * z_pitching +
      0.25 * z_command +
      0.25 * z_miss +
      0.20 * z_contact
  ) %>%

  # ==========================================================
  # CREATE PERCENTILES
  # ==========================================================

  dplyr::mutate(
    dplyr::across(
      dplyr::starts_with("elite_"),
      ~ dplyr::percent_rank(.x),
      .names = "{.col}_percentile"
    )
  ) %>%

  # ==========================================================
  # IDENTIFY TOP-10% PITCHERS
  # ==========================================================

  dplyr::mutate(
    dplyr::across(
      dplyr::ends_with("_percentile"),
      ~ dplyr::if_else(
        is.na(.x),
        NA,
        .x >= 0.90
      ),
      .names = "{.col}_top_10"
    )
  )

# ============================================================
# VIEW RESULTS
# ============================================================

elite_test %>%
  dplyr::select(
    pitcher,
    player_full_name,
    stuff_plus,
    location_plus,
    pitching_plus,
    command_score,
    miss_score,
    contact_score,
    dplyr::starts_with("elite_")
  ) %>%
  dplyr::arrange(
    dplyr::desc(elite_original)
  ) %>%
  View()
```

## Validation Outcomes
```{r run value}
run_value_tbl <- PBP %>%
  filter(!is.na(pitcher)) %>%
  group_by(pitcher) %>%
  summarise(
    total_delta_pitcher_run_exp =
      sum(delta_pitcher_run_exp, na.rm = TRUE),

    n_pitches = n(),

    drx100 =
      100 * total_delta_pitcher_run_exp / n_pitches,

    .groups = "drop"
  )
```

```{r bench table}
bench_tbl <- P_stats %>%
  transmute(
    pitcher,
    ERA = era,
    IP = innings_pitched,
    K9 = strikeouts_per9inn,
    BB9 = walks_per9inn,
    HR9 = home_runs_per9,
    WHIP = whip
  ) %>%
  distinct(pitcher, .keep_all = TRUE)
```

``` {r fip components}
fip_components <- PBP %>%
  filter(!is.na(pitcher)) %>%
  group_by(pitcher) %>%
  summarise(
    HR = sum(events == "home_run", na.rm = TRUE),

    BB = sum(
      events %in% c("walk", "intent_walk"),
      na.rm = TRUE
    ),

    HBP = sum(
      events == "hit_by_pitch",
      na.rm = TRUE
    ),

    K = sum(
      events %in% c(
        "strikeout",
        "strikeout_double_play"
      ),
      na.rm = TRUE
    ),

    .groups = "drop"
  )
```

```{r lg_const}
lg_const <- 3.10

fip_tbl <- fip_components %>%
  dplyr::left_join(
    bench_tbl %>%
      dplyr::select(pitcher, IP),
    by = "pitcher"
  ) %>%
  dplyr::filter(!is.na(IP), IP > 0) %>%
  dplyr::mutate(
    FIP_est =
      (13 * HR + 3 * (BB + HBP) - 2 * K) / IP +
      lg_const
  ) %>%
  dplyr::select(
    pitcher,
    FIP_est
  )
```

```{r val_dat}
val_dat <- elite_test %>%
  left_join(run_value_tbl, by = "pitcher") %>%
  left_join(bench_tbl, by = "pitcher") %>%
  left_join(fip_tbl, by = "pitcher")
```

```{r variables}
model_variables <- c(
  "elite_original",
  "elite_pitching_only",
  "elite_stuff_location",
  "elite_all_three",
  "z_stuff",
  "z_location",
  "z_pitching",
  "z_command",
  "z_miss",
  "z_contact",
  "drx100"
)

val_common <- val_dat %>%
  filter(
    if_all(
      all_of(model_variables),
      ~ !is.na(.x)
    )
  )
```

```{r}
cat("Common validation sample:", nrow(val_common), "pitchers")
```

## direction of drx100
```{r val_common}
val_common %>%
  summarise(
    cor_drx_ERA = cor(drx100, ERA, use = "complete.obs"),
    cor_drx_FIP = cor(drx100, FIP_est, use = "complete.obs"),
    cor_drx_WHIP = cor(drx100, WHIP, use = "complete.obs")
  )
```

## Correlation & redundancy analysis
```{r}
skill_vars <- val_common %>%
  dplyr::select(
    stuff_plus,
    location_plus,
    pitching_plus,
    command_score,
    miss_score,
    contact_score
  )

skill_cor <- cor(
  skill_vars,
  use = "pairwise.complete.obs",
  method = "pearson"
)

round(skill_cor, 3)

corrplot::corrplot(
  skill_cor,
  method = "number"
)
```
## Spearman Correlation
```{r spearman}
skill_cor_spearman <- cor(
  skill_vars,
  use = "pairwise.complete.obs",
  method = "spearman"
)

round(skill_cor_spearman, 3)
```
## Construct-Validity tests
```{r construct-validity}
construct_models <- list(
  stuff_predicts_miss =
    lm(z_miss ~ z_stuff, data = val_common),

  location_predicts_command =
    lm(z_command ~ z_location, data = val_common),

  stuff_predicts_K9 =
    lm(K9 ~ z_stuff, data = val_common),

  location_predicts_BB9 =
    lm(BB9 ~ z_location, data = val_common),

  pitching_predicts_drx =
    lm(drx100 ~ z_pitching, data = val_common)
)

construct_results <- imap_dfr(
  construct_models,
  ~ tidy(.x, conf.int = TRUE) %>%
    mutate(model = .y)
)

construct_results
```

## Compare each elite index against drx100
```{r}
index_names <- c(
  "elite_original",
  "elite_pitching_only",
  "elite_stuff_location",
  "elite_all_three"
)

index_models <- set_names(index_names) %>%
  map(
    ~ lm(
      reformulate(.x, response = "drx100"),
      data = val_common
    )
  )
```

```{r}
index_fit_tbl <- purrr::imap_dfr(
  index_models,
  function(model, model_name) {
    broom::glance(model) %>%
      dplyr::mutate(model = model_name)
  }
) %>%
  dplyr::select(
    model,
    nobs,
    r.squared,
    adj.r.squared,
    sigma,
    AIC,
    BIC
  ) %>%
  dplyr::arrange(BIC)

knitr::kable(index_fit_tbl, digits = 3)
```

```{r}
index_coef_tbl <- imap_dfr(
  index_models,
  function(model, model_name) {
    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(model = model_name)
  }
)

kable(index_coef_tbl, digits = 3)
```

## Rax Index Correlations
```{r}
index_correlations <- map_dfr(
  index_names,
  function(index_name) {
    tibble(
      model = index_name,

      pearson =
        cor(
          val_common[[index_name]],
          val_common$drx100,
          use = "complete.obs",
          method = "pearson"
        ),

      spearman =
        cor(
          val_common[[index_name]],
          val_common$drx100,
          use = "complete.obs",
          method = "spearman"
        )
    )
  }
)

kable(index_correlations, digits = 3)
```

## Component-level models
```{r }
component_models <- list(
  original_components =
    lm(
      drx100 ~
        z_stuff +
        z_command +
        z_miss +
        z_contact,
      data = val_common
    ),

  pitching_components =
    lm(
      drx100 ~
        z_pitching +
        z_command +
        z_miss +
        z_contact,
      data = val_common
    ),

  stuff_location_components =
    lm(
      drx100 ~
        z_stuff +
        z_location +
        z_command +
        z_miss +
        z_contact,
      data = val_common
    ),

  all_three_components =
    lm(
      drx100 ~
        z_stuff +
        z_location +
        z_pitching +
        z_command +
        z_miss +
        z_contact,
      data = val_common
    )
)
```

```{r }
component_fit_tbl <- purrr::imap_dfr(
  component_models,
  function(model, model_name) {
    broom::glance(model) %>%
      dplyr::mutate(model = model_name)
  }
) %>%
  dplyr::select(
    model,
    nobs,
    r.squared,
    adj.r.squared,
    sigma,
    AIC,
    BIC
  ) %>%
  dplyr::arrange(BIC)

knitr::kable(
  component_fit_tbl,
  digits = 3
)
```
## Nested tests for incremental info
```{r}
m_base <- lm(
  drx100 ~
    z_command +
    z_miss +
    z_contact,
  data = val_common
)

m_add_stuff <- update(
  m_base,
  . ~ . + z_stuff
)

m_add_stuff_location <- update(
  m_base,
  . ~ . + z_stuff + z_location
)

m_add_pitching <- update(
  m_base,
  . ~ . + z_pitching
)

m_add_all <- update(
  m_base,
  . ~ . + z_stuff + z_location + z_pitching
)
```

```{r}
anova(m_base, m_add_stuff)
anova(m_base, m_add_stuff_location)
anova(m_base, m_add_pitching)
anova(m_base, m_add_all)
anova(m_add_stuff_location, m_add_all)
```
## Multicollinearity
```{r}
vif_results <- imap_dfr(
  component_models,
  function(model, model_name) {
    vif_values <- car::vif(model)

    tibble(
      model = model_name,
      variable = names(vif_values),
      VIF = as.numeric(vif_values)
    )
  }
)

kable(vif_results, digits = 2)
```
## Benchmark Comparison
```{r}
benchmark_variables <- c(
  "ERA",
  "FIP_est",
  "WHIP",
  "K9",
  "BB9",
  "HR9"
)

benchmark_variables <- benchmark_variables[
  benchmark_variables %in% names(val_common)
]
```

```{r}
single_predictors <- c(
  index_names,
  "stuff_plus",
  "location_plus",
  "pitching_plus",
  benchmark_variables
)

single_metric_models <- set_names(single_predictors) %>%
  map(
    ~ lm(
      reformulate(.x, response = "drx100"),
      data = val_common
    )
  )
```

```{r}
single_metric_fit <- purrr::imap_dfr(
  single_metric_models,
  function(model, model_name) {
    broom::glance(model) %>%
      dplyr::mutate(model = model_name)
  }
) %>%
  dplyr::select(
    model,
    nobs,
    r.squared,
    adj.r.squared,
    sigma,
    AIC,
    BIC
  ) %>%
  dplyr::arrange(BIC)

knitr::kable(
  single_metric_fit,
  digits = 3
)
```


## Incremental Value beyond benchmarks
```{r}
incremental_models <- crossing(
  index = index_names,
  benchmark = intersect(
    c("ERA", "FIP_est", "WHIP"),
    names(val_common)
  )
) %>%
  mutate(
    benchmark_only = map(
      benchmark,
      ~ lm(
        reformulate(.x, response = "drx100"),
        data = val_common
      )
    ),

    benchmark_plus_index = map2(
      benchmark,
      index,
      ~ lm(
        reformulate(
          c(.x, .y),
          response = "drx100"
        ),
        data = val_common
      )
    ),

    comparison = map2(
      benchmark_only,
      benchmark_plus_index,
      anova
    ),

    base_adj_r2 = map_dbl(
      benchmark_only,
      ~ glance(.x)$adj.r.squared
    ),

    combined_adj_r2 = map_dbl(
      benchmark_plus_index,
      ~ glance(.x)$adj.r.squared
    ),

    delta_adj_r2 =
      combined_adj_r2 - base_adj_r2,

    combined_AIC = map_dbl(
      benchmark_plus_index,
      AIC
    )
  )
```

```{r}
incremental_models %>%
  dplyr::select(
    index,
    benchmark,
    base_adj_r2,
    combined_adj_r2,
    delta_adj_r2,
    combined_AIC
  ) %>%
  dplyr::arrange(
    dplyr::desc(delta_adj_r2)
  ) %>%
  knitr::kable(
    digits = 3
  )
```

## Repeated cross-validation
```{r}
repeated_kfold_lm <- function(
  data,
  formula,
  K = 10,
  repeats = 25,
  seed = 42
) {

  set.seed(seed)

  results <- map_dfr(
    seq_len(repeats),
    function(repeat_id) {

      fold_id <- sample(
        rep(1:K, length.out = nrow(data))
      )

      predictions <- rep(NA_real_, nrow(data))

      for (k in seq_len(K)) {

        train <- data[
          fold_id != k,
          ,
          drop = FALSE
        ]

        test <- data[
          fold_id == k,
          ,
          drop = FALSE
        ]

        fit <- lm(formula, data = train)

        predictions[fold_id == k] <-
          predict(fit, newdata = test)
      }

      observed <- model.response(
        model.frame(formula, data = data)
      )

      tibble(
        repeat_id = repeat_id,

        RMSE = sqrt(
          mean(
            (observed - predictions)^2,
            na.rm = TRUE
          )
        ),

        R2 = 1 -
          sum(
            (observed - predictions)^2,
            na.rm = TRUE
          ) /
          sum(
            (observed - mean(observed))^2,
            na.rm = TRUE
          ),

        correlation = cor(
          observed,
          predictions,
          use = "complete.obs"
        )
      )
    }
  )

  results
}
```

```{r}
cv_formulas <- list(
  elite_original =
    drx100 ~ elite_original,

  elite_pitching_only =
    drx100 ~ elite_pitching_only,

  elite_stuff_location =
    drx100 ~ elite_stuff_location,

  elite_all_three =
    drx100 ~ elite_all_three,

  original_components =
    drx100 ~
      z_stuff +
      z_command +
      z_miss +
      z_contact,

  pitching_components =
    drx100 ~
      z_pitching +
      z_command +
      z_miss +
      z_contact,

  stuff_location_components =
    drx100 ~
      z_stuff +
      z_location +
      z_command +
      z_miss +
      z_contact,

  all_three_components =
    drx100 ~
      z_stuff +
      z_location +
      z_pitching +
      z_command +
      z_miss +
      z_contact
)
```

```{r}
cv_results <- imap_dfr(
  cv_formulas,
  function(formula, model_name) {
    repeated_kfold_lm(
      data = val_common,
      formula = formula,
      K = 10,
      repeats = 25
    ) %>%
      mutate(model = model_name)
  }
)
```

```{r}
cv_summary <- cv_results %>%
  group_by(model) %>%
  summarise(
    mean_RMSE = mean(RMSE),
    sd_RMSE = sd(RMSE),
    mean_R2 = mean(R2),
    sd_R2 = sd(R2),
    mean_correlation = mean(correlation),
    sd_correlation = sd(correlation),
    .groups = "drop"
  ) %>%
  arrange(mean_RMSE)

kable(cv_summary, digits = 3)
```

## Ranking correlations
```{r}
ranking_cor <- val_common %>%
  dplyr::select(
    dplyr::all_of(index_names)
  ) %>%
  cor(
    use = "complete.obs",
    method = "spearman"
  )

round(ranking_cor, 3)

corrplot::corrplot(
  ranking_cor,
  method = "number"
)
```

## Top-10% overlap
```{r}
top10_sets <- map(
  index_names,
  function(index_name) {
    val_common %>%
      filter(
        percent_rank(.data[[index_name]]) >= 0.90
      ) %>%
      pull(pitcher)
  }
)

names(top10_sets) <- index_names
```

```{r}
jaccard_similarity <- function(x, y) {
  length(intersect(x, y)) /
    length(union(x, y))
}
```

```{r}
top10_overlap <- crossing(
  model_1 = index_names,
  model_2 = index_names
) %>%
  mutate(
    jaccard = map2_dbl(
      model_1,
      model_2,
      ~ jaccard_similarity(
        top10_sets[[.x]],
        top10_sets[[.y]]
      )
    )
  )

top10_overlap
```

## Largest ranking changes
```{r}
rank_change_tbl <- val_common %>%
  transmute(
    pitcher,
    player_full_name,

    rank_original =
      min_rank(desc(elite_original)),

    rank_pitching_only =
      min_rank(desc(elite_pitching_only)),

    rank_stuff_location =
      min_rank(desc(elite_stuff_location)),

    rank_all_three =
      min_rank(desc(elite_all_three)),

    original_to_pitching_change =
      rank_original - rank_pitching_only,

    original_to_stuff_location_change =
      rank_original - rank_stuff_location,

    original_to_all_three_change =
      rank_original - rank_all_three
  )
```

```{r}
rank_change_tbl %>%
  arrange(
    desc(abs(original_to_pitching_change))
  ) %>%
  slice_head(n = 20) %>%
  kable()
```

## Elite threshold sensitivity
```{r}
index_names <- c(
  "elite_original",
  "elite_pitching_only",
  "elite_stuff_location",
  "elite_all_three"
)

thresholds <- c(
  0.80,
  0.85,
  0.90,
  0.95
)

threshold_results <- tidyr::crossing(
  index = index_names,
  threshold = thresholds
) %>%
  dplyr::mutate(
    results = purrr::map2(
      index,
      threshold,
      function(index_name, cutoff) {

        temp <- val_common

        temp$percentile <- dplyr::percent_rank(
          temp[[index_name]]
        )

        temp$elite_group <-
          temp$percentile >= cutoff

        temp %>%
          dplyr::group_by(elite_group) %>%
          dplyr::summarise(
            n = dplyr::n(),
            mean_drx100 = mean(
              drx100,
              na.rm = TRUE
            ),
            median_drx100 = median(
              drx100,
              na.rm = TRUE
            ),
            .groups = "drop"
          )
      }
    )
  ) %>%
  tidyr::unnest(
    cols = results
  )

threshold_results %>%
  dplyr::arrange(
    index,
    threshold,
    elite_group
  ) %>%
  knitr::kable(
    digits = 3,
    caption = "Elite Threshold Sensitivity Results"
  )
```

## Sample- size sensitivity
```{r}
pitch_thresholds <- c(
  250,
  500,
  750,
  1000
)

sample_sensitivity <- tidyr::crossing(
  model = index_names,
  minimum_pitches = pitch_thresholds
) %>%
  dplyr::mutate(
    fit = purrr::map2(
      model,
      minimum_pitches,
      function(model_name, pitch_cutoff) {

        sample_dat <- val_common %>%
          dplyr::filter(
            n_pitches >= pitch_cutoff
          )

        if (nrow(sample_dat) < 20) {
          return(NULL)
        }

        stats::lm(
          stats::reformulate(
            model_name,
            response = "drx100"
          ),
          data = sample_dat
        )
      }
    ),

    results = purrr::map(
      fit,
      function(model_object) {

        if (is.null(model_object)) {
          return(
            tibble::tibble(
              n = NA_integer_,
              estimate = NA_real_,
              r_squared = NA_real_,
              adj_r_squared = NA_real_,
              RMSE = NA_real_
            )
          )
        }

        coefficient <- broom::tidy(
          model_object
        ) %>%
          dplyr::filter(
            term != "(Intercept)"
          ) %>%
          dplyr::pull(
            estimate
          )

        fit_stats <- broom::glance(
          model_object
        )

        tibble::tibble(
          n = as.integer(fit_stats$nobs),
          estimate = coefficient,
          r_squared = fit_stats$r.squared,
          adj_r_squared = fit_stats$adj.r.squared,
          RMSE = fit_stats$sigma
        )
      }
    )
  ) %>%
  dplyr::select(
    -fit
  ) %>%
  tidyr::unnest(
    cols = results
  ) %>%
  dplyr::arrange(
    minimum_pitches,
    dplyr::desc(adj_r_squared)
  )

knitr::kable(
  sample_sensitivity,
  digits = 3
)
```

## Outlier & Influence Analysis
```{r}
preferred_candidate <- stats::lm(
  drx100 ~ elite_stuff_location,
  data = val_common
)

influence_tbl <- broom::augment(
  preferred_candidate
) %>%
  dplyr::bind_cols(
    val_common %>%
      dplyr::select(
        pitcher,
        player_full_name
      )
  ) %>%
  dplyr::arrange(
    dplyr::desc(.cooksd)
  )

influence_tbl %>%
  dplyr::select(
    pitcher,
    player_full_name,
    .fitted,
    .resid,
    .std.resid,
    .hat,
    .cooksd
  ) %>%
  dplyr::slice_head(
    n = 15
  ) %>%
  knitr::kable(
    digits = 3
  )
```

```{r}
cooks_cutoff <- 4 / nrow(val_common)

influential_pitchers <- influence_tbl %>%
  filter(.cooksd > cooks_cutoff)

influential_pitchers
```

```{r}
robust_sample <- val_common %>%
  filter(
    !pitcher %in% influential_pitchers$pitcher
  )

lm(
  drx100 ~ elite_stuff_location,
  data = robust_sample
) %>%
  summary()
```
## Robust regression
```{r}
robust_models <- map(
  index_names,
  function(index_name) {
    MASS::rlm(
      reformulate(
        index_name,
        response = "drx100"
      ),
      data = val_common
    )
  }
)

names(robust_models) <- index_names
```

```{r}
robust_coef_tbl <- imap_dfr(
  robust_models,
  function(model, model_name) {
    tidy(model) %>%
      mutate(model = model_name)
  }
)

robust_coef_tbl
```

## Weight sensitivity
```{r}
weight_scenarios <- tribble(
  ~scenario, ~fg_weight, ~command_weight, ~miss_weight, ~contact_weight,
  "FG 20",       0.20,          0.30,        0.30,          0.20,
  "FG 25",       0.25,          0.275,       0.275,         0.20,
  "FG 30",       0.30,          0.25,        0.25,          0.20,
  "FG 35",       0.35,          0.225,       0.225,         0.20,
  "Equal",       0.25,          0.25,        0.25,          0.25
)
```

```{r}
weight_sensitivity <- weight_scenarios %>%
  dplyr::mutate(
    results = purrr::pmap(
      list(
        scenario,
        fg_weight,
        command_weight,
        miss_weight,
        contact_weight
      ),
      function(
        scenario,
        fg_weight,
        command_weight,
        miss_weight,
        contact_weight
      ) {

        temp <- val_common %>%
          dplyr::mutate(
            test_index =
              fg_weight * z_pitching +
              command_weight * z_command +
              miss_weight * z_miss +
              contact_weight * z_contact
          )

        fit <- stats::lm(
          drx100 ~ test_index,
          data = temp
        )

        tibble::tibble(
          scenario = scenario,

          spearman_drx = stats::cor(
            temp$test_index,
            temp$drx100,
            method = "spearman",
            use = "complete.obs"
          ),

          estimate = broom::tidy(fit) %>%
            dplyr::filter(
              term == "test_index"
            ) %>%
            dplyr::pull(
              estimate
            ),

          adj_r_squared = broom::glance(fit)$adj.r.squared,

          RMSE = broom::glance(fit)$sigma
        )
      }
    )
  ) %>%
  dplyr::select(
    results
  ) %>%
  tidyr::unnest(
    cols = results
  )

weight_sensitivity
```

## Bootstrap model comparison
```{r}
bootstrap_compare <- function(
  data,
  index_names,
  B = 1000,
  seed = 42
) {

  set.seed(seed)

  map_dfr(
    seq_len(B),
    function(b) {

      boot_dat <- data[
        sample(
          seq_len(nrow(data)),
          replace = TRUE
        ),
        ,
        drop = FALSE
      ]

      map_dfr(
        index_names,
        function(index_name) {

          fit <- lm(
            reformulate(
              index_name,
              response = "drx100"
            ),
            data = boot_dat
          )

          tibble(
            bootstrap = b,
            model = index_name,
            estimate =
              coef(fit)[[index_name]],
            adj_r_squared =
              glance(fit)$adj.r.squared,
            RMSE =
              glance(fit)$sigma
          )
        }
      )
    }
  )
}
```

```{r}
bootstrap_results <- bootstrap_compare(
  data = val_common,
  index_names = index_names,
  B = 1000
)
```

```{r}
bootstrap_summary <- bootstrap_results %>%
  group_by(model) %>%
  summarise(
    mean_estimate = mean(estimate, na.rm = TRUE),
    estimate_low = quantile(
      estimate,
      0.025,
      na.rm = TRUE
    ),
    estimate_high = quantile(
      estimate,
      0.975,
      na.rm = TRUE
    ),
    mean_adj_r2 = mean(
      adj_r_squared,
      na.rm = TRUE
    ),
    mean_RMSE = mean(
      RMSE,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  arrange(mean_RMSE)

kable(bootstrap_summary, digits = 3)
```

## Final comparison table
```{r}
final_comparison <- index_fit_tbl %>%
  dplyr::rename(
    in_sample_R2 = r.squared,
    in_sample_adj_R2 = adj.r.squared,
    in_sample_RMSE = sigma
  ) %>%

  dplyr::left_join(
    cv_summary %>%
      dplyr::select(
        model,
        CV_RMSE = mean_RMSE,
        CV_R2 = mean_R2,
        CV_correlation = mean_correlation
      ),
    by = "model"
  ) %>%

  dplyr::left_join(
    index_correlations %>%
      dplyr::rename(
        outcome_pearson = pearson,
        outcome_spearman = spearman
      ),
    by = "model"
  ) %>%

  dplyr::left_join(
    bootstrap_summary %>%
      dplyr::select(
        model,
        bootstrap_adj_R2 = mean_adj_r2,
        bootstrap_RMSE = mean_RMSE,
        estimate_low,
        estimate_high
      ),
    by = "model"
  ) %>%

  dplyr::arrange(
    CV_RMSE
  )

knitr::kable(
  final_comparison,
  digits = 3
)
```

