---
title: "Project 2"
author: "Hannah Dean"
date: "2025-10-24"
output:
  word_document:
    toc: true
    toc_depth: '3'
  html_document:
    theme:
      version: 5
      bootswatch: cosmo
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: hide
    df_print: paged
    fig_caption: true
    highlight: tango
    smooth_scroll: true
editor_options:
  markdown:
    wrap: 72
knitr:
  opts_chunk:
    echo: false
    message: false
    warning: false
---

## Design & Simulation

#### Topic & Rationale

Spin Rate Tiers Effect on Exit Velocity

Does exit velocity differ across spin rater tiers (Low, Medium, High)
and between batter handedness (Left vs Right)? Does the effect of spin
rate on exit velocity depend on batter handiness?

Main Effect Spin Rate Tier

-   **H₀₁:** \$ \mu\*{Low} =\* \mu{Med} = \mu\_{High} \$\
    Exit velocity does not differ across spin-rate tiers.\
-   **H₁₁:** At least one tier mean differs (\$ \mu\_i \ne \mu\_j \$ for
    some *i*, *j*).\
    Spin rate tier has a significant effect on exit velocity.

Main Effect Batter Handiness

-   **H₀₂:** \$ \mu\*{Left} =\* \mu{Right} \$\
    Batter handedness does not influence exit velocity.\
-   **H₁₂:** \$ \mu\*{Left}\* \ne \mu{Right} \$\
    Left- and right-handed hitters differ in average exit velocity.

Internationale Effect (Spin Rate Tiers x Batter Handiness)

-   **H₀₃:** There is no interaction between spin-rate tier and batter
    handedness.\
    The effect of spin rate on exit velocity is the same for both left-
    and right-handed batters.\
-   **H₁₃:** There is a significant interaction.\
    The relationship between spin rate and exit velocity changes
    depending on batter handedness.

#### Factor Definition

Factor A: (Spin Rate Tiers) by Quantile • Low SR • Medium SR • High SR

Factor B: (Batter Handiness) • Left • Right

All 6 treatments: LSRleft, LSRright, MSRleft, MSRright, HSRleft,
HSRright

#### Data Construction

```{r Data Construction}
# ---- Packages ----
library(vroom)
library(dplyr)
library(stringr)
library(lubridate)
library(purrr)
library(tibble)
library(readr)
library(tidyr)
library(multcomp)
library(ggplot2)

# "C:\Users\hanna\OneDrive\Documents\Pionneer League Games 2025"
#"C:\\Users\\hanna\\OneDrive\\Documents\\Pioneer League Games 2025"

# ---- Find all official games in "Pioneer League Games 2025" File ----
#base_dir <- (choose.dir(caption = "Select the 'Pioneer League Games 2025' folder"))

base_dir <- ("C:\\Users\\hanna\\OneDrive\\Documents\\Pionneer League Games 2025")

list_official_csvs <- function(base_dir, strict_structure = TRUE) {
  all_csvs <- list.files(base_dir, pattern = "\\.csv$", recursive = TRUE, full.names = TRUE, ignore.case = TRUE)
  cat("Total .csv found (anywhere under base):", length(all_csvs), "\n")
  
  # Must be inside .../(5..9 or 05..09)/(day)/CSV/...
  if (strict_structure) {
    under_csv <- grepl("[/\\\\](0?[5-9])[/\\\\]\\d{1,2}[/\\\\]CSV[/\\\\]", all_csvs, ignore.case = TRUE)
  } else {
    # Relaxed: anywhere there is a CSV folder in the path
    under_csv <- grepl("[/\\\\]CSV[/\\\\]", all_csvs, ignore.case = TRUE)
  }
  cat("…in CSV folders with month/day structure:", sum(under_csv), "\n")
  
  # Official filename: YYYYMMDD-Field-Game#.csv  (case-insensitive)
  fname <- basename(all_csvs)
  looks_official <- grepl("^\\d{8}-[^-]+-\\d+\\.csv$", fname, ignore.case = TRUE)
  cat("…matching official filename pattern:", sum(looks_official), "\n")
  
  # Exclude any *_unverified* variants in "CSV" folder
  not_unverified <- !grepl("_unverified", fname, ignore.case = TRUE)
  cat("…excluding _unverified:", sum(not_unverified), "\n")
  
  keep <- under_csv & looks_official & not_unverified
  files <- sort(all_csvs[keep])
  cat("Official files kept:", length(files), "\n")
  
  if (length(files) == 0) {
    cat("\nNo matches. Quick hints:\n",
        "- Check that month folders are 5..9 or 05..09.\n",
        "- Ensure there is a 'CSV' folder (exact name, any case) at the 3rd level.\n",
        "- Confirm filenames look like 20250902-DehlerPark-1.csv (no spaces, hyphens exactly).\n",
        "- If your structure differs, set strict_structure = FALSE.\n", sep = "")
  }
  files
}

# ---- Run the finder ----
official_files <- list_official_csvs(base_dir, strict_structure = TRUE)
# If you get 0, try relaxed search:
if (length(official_files) == 0) {
  cat("\nRetrying with relaxed structure (any 'CSV' folder)…\n")
  official_files <- list_official_csvs(base_dir, strict_structure = FALSE)
}

# ---- Balls in Play ----
# We just want rows where there is a ball in play so create new spreadsheet with 
# value: "InPlay" in variable:"PitchCall"

pbl_official <- vroom::vroom(
  official_files,
  .name_repair = "unique",
  altrep = TRUE,
  id = "source_file",
  progress = TRUE
) %>%
  mutate(
    file_name     = basename(source_file),
    game_date_chr = str_match(file_name, "^(\\d{8})-")[,2],
    field         = str_match(file_name, "^\\d{8}-([^-]+)-\\d+\\.csv$")[,2],
    game_no       = suppressWarnings(as.integer(str_match(file_name, "-(\\d+)\\.csv$")[,2])),
    game_date     = suppressWarnings(lubridate::ymd(game_date_chr))
  ) %>%
  dplyr::select(source_file, file_name, game_date, field, game_no, dplyr::everything())


# column grabber for "PitchCall"
col_like <- function(df, target) {
  nm <- names(df)
  hit <- which(tolower(gsub("\\s+", "", nm)) == tolower(gsub("\\s+", "", target)))
  if (length(hit) == 0) return(NA_character_)
  nm[hit[1]]
}
pc_col <- col_like(pbl_official, "PitchCall")
if (is.na(pc_col)) stop("Couldn't find a 'PitchCall' column in pbl_official.")

# Filter InPlay and write CSV
out_dir <- file.path(base_dir, "Outputs")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

all_inplay <- pbl_official %>%
  filter(!is.na(.data[[pc_col]])) %>%
  mutate(!!pc_col := trimws(.data[[pc_col]])) %>%
  filter(.data[[pc_col]] == "InPlay")

out_path <- file.path(out_dir, "PioneerLeague_InPlay_AllRows.csv")
readr::write_csv(all_inplay, out_path)

cat("Saved", nrow(all_inplay), "rows to:\n", out_path, "\n")

# ---- Edited Variables ----
# Edit Data so it only has Variables needed for research

# Required Variables
required <- c(
  "BatterSide","PitcherThrows","SpinRate","AutoPitchType","ExitSpeed","Angle",
  "RelSpeed","ZoneSpeed","PlateLocHeight","PlateLocSide","AutoHitType",
  "Direction","HitSpinRate"
)

# Make the trimmed tibble in memory
stopifnot(exists("all_inplay"))
missing <- setdiff(required, names(all_inplay))
if (length(missing)) stop(paste("Missing columns in `all_inplay`:", paste(missing, collapse=", ")))

inplay_var <- dplyr::select(all_inplay, all_of(required))

# Quick peek
glimpse(inplay_var)
nrow(inplay_var); ncol(inplay_var); names(inplay_var)

# (Optional) Save it too
out_dir <- file.path(base_dir, "Outputs")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
out_path <- file.path(out_dir, "PioneerLeague_InPlay_SelectedCols_STRICT.csv")
write_csv(inplay_var, out_path)
cat("Saved:", out_path, "\n")

# ---- Delete NA's ----
# Get rid of rows with NA's

inplay_var_clean <- inplay_var %>%
  # 1) Trim and convert "" -> NA ONLY for character columns
  mutate(across(where(is.character), ~ na_if(trimws(.), ""))) %>%
  # 2) (Optional) Parse obvious numeric columns that might have come in as text
  mutate(across(c(SpinRate, ExitSpeed, Angle, RelSpeed, ZoneSpeed,
                  PlateLocHeight, PlateLocSide, HitSpinRate, Direction),
                ~ suppressWarnings(as.numeric(.)), .names = "{.col}")) %>%
  # 3) Drop rows with any NA in any column
  drop_na()

# Quick sanity check + save
removed <- nrow(inplay_var) - nrow(inplay_var_clean)
cat("Rows before:", nrow(inplay_var),
    " | after:", nrow(inplay_var_clean),
    " | removed:", removed, "\n")

out_dir <- file.path(base_dir, "Outputs")
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
out_path <- file.path(out_dir, "PioneerLeague_InPlay_SelectedCols_STRICT_noNA.csv")
write_csv(inplay_var_clean, out_path)
cat("Saved cleaned file to:\n", out_path, "\n")

# ---- Spin tirers & Handiness ----
stopifnot(exists("inplay_var_clean"))

df <- inplay_var_clean %>%
  filter(!is.na(SpinRate), !is.na(ExitSpeed), !is.na(BatterSide)) %>%
  mutate(
    SpinTier = cut(
      SpinRate,
      breaks = quantile(SpinRate, probs = c(0, 1/3, 2/3, 1), na.rm = TRUE),
      labels = c("Low", "Medium", "High"),
      include.lowest = TRUE
    ),
    BatterSide = case_when(
      grepl("^l|left",  tolower(BatterSide)) ~ "Left",
      grepl("^r|right", tolower(BatterSide)) ~ "Right",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(BatterSide)) %>%
  mutate(
    SpinTier   = factor(SpinTier, levels = c("Low","Medium","High")),
    BatterSide = factor(BatterSide, levels = c("Left","Right")),
    Combo = interaction(SpinTier, BatterSide, sep = "_")   # 6 combined groups
  ) %>%
  droplevels()

# ---- Tibble Table ----
summary_tibble <- df %>%
  count(SpinTier, BatterSide) %>%
  pivot_wider(
    names_from = SpinTier,
    values_from = n,
    values_fill = 0
  ) %>%
  rename(
    Batter_Side = BatterSide
  ) %>%
  arrange(Batter_Side)

# Print nicely
summary_tibble
```

## Parameter Estimation

#### Model Fitting & Output

```{r Model Fit Statistics}
lm1  <- lm(ExitSpeed ~ Combo, data = df)
summary(lm1)
```

No single groups mean Exit Velocity differs significantly Residual
standard error is 13.5 mph which is a big error for exit velocity
meaning there inst much variation R\^2 is equal to .06% meaning that
only .06%is explained by the different groups The F-statistic shows us
that there may be some differences between the groups exit velocity but
there aren't any strong and significant differences that are meaningful

```{r ANOVA Table}
fit1 <- aov(ExitSpeed ~ Combo, data = df)
summary(fit1)
```

Sum of Squares also shows us that the variables explain a tiny fraction
of the total variation The P value of the F statistic shows that there
is a statistically significant effect on at least one of the groups
However, despite statistical significance the effect size is very small
and effects .1% of the variation in Exit Velocity

#### Parameter Interpertation

$$
Y_{ij} = \mu + \alpha_i + \epsilon_{ij}
$$ where\
- $Y_{ij}$: Exit velocity for the *jth* observation in group *i*\
- $\mu$: Overall (grand) mean\
- $\alpha_i$: Effect of the *ith* Combo group (deviation from grand
mean)\
- $\epsilon_{ij}$: Random error term, assumed $\sim N(0, \sigma^2)$

$$
ExitSpeed_{ij} = \mu + \alpha_{\text{Combo}_i} + \epsilon_{ij}
$$

#### Group Summary Table

```{r Mean & SDs}
group_summary <- df %>%
  group_by(Combo) %>%
  summarise(
    n = n(),
    mean_ES = mean(ExitSpeed),
    sd_ES = sd(ExitSpeed),
    .groups = "drop"
  )
print(group_summary)

overall_mean <- mean(df$ExitSpeed)
alpha_hats <- group_summary %>%
  mutate(alpha_hat = mean_ES - overall_mean)

print(alpha_hats)
```

The mean Exit Speed is around 89mph for each group with a standard
deviation of 12-14 mph which are larger standards of deviation for the
data set, The low_right group had the largest exit velocity of 89.95 mph
The high_left group had the lowest exit velocity of 88.84 mph So the
range between the lowest and the largest exit velocity is 1.1mph which
is a difference but it is not a ver big difference in terms of the
context

#### ANOVA Table & Hypotheses

F-Stat: F(5, 19892) = 2.44 P-Value: .032 Significance level: .05

#### F-test Interpertation

since p = .032 \< .05, we reject the null hypothesis meaning that there
is a significant difference in mean Exit Velocity among the groups.

That being said the statistically significant differences is very small
roughly 1mph and it isn't very meaningful in terms of the context and
the groups have a negligible influence on Exit Velocity.

## Contrasts & Post-Hoc Test

```{r Contrast Prep}
# Contrast Matrix
K_combo <- rbind(
  # High vs (Low+Med)/2 within each side
  High_vs_Others_Left  = c( Low_Left = -0.5, Medium_Left = -0.5, High_Left = 1,
                            Low_Right = 0,   Medium_Right = 0,   High_Right = 0 ),
  High_vs_Others_Right = c( Low_Left = 0,   Medium_Left = 0,   High_Left = 0,
                            Low_Right = -0.5, Medium_Right = -0.5, High_Right = 1 ),

  # Low vs Medium within each side
  Low_vs_Med_Left  = c( Low_Left = 1,  Medium_Left = -1, High_Left = 0,
                        Low_Right = 0, Medium_Right = 0,  High_Right = 0 ),
  Low_vs_Med_Right = c( Low_Left = 0, Medium_Left = 0,  High_Left = 0,
                        Low_Right = 1, Medium_Right = -1, High_Right = 0 ),

  # Left vs Right within each spin tier
  Left_vs_Right_Low   = c( Low_Left = -1, Medium_Left = 0, High_Left = 0,
                           Low_Right = 1,  Medium_Right = 0, High_Right = 0 ),
  Left_vs_Right_Med   = c( Low_Left = 0, Medium_Left = -1, High_Left = 0,
                           Low_Right = 0,  Medium_Right = 1,  High_Right = 0 ),
  Left_vs_Right_High  = c( Low_Left = 0, Medium_Left = 0, High_Left = -1,
                           Low_Right = 0,  Medium_Right = 0, High_Right = 1 )
)


# ---- Contrasts ----
# Add the two contrasts
K_combo <- rbind(
  K_combo,

  # C1: High spin (avg of High_Left, High_Right) vs Low spin (avg of Low_Left, Low_Right)
  High_vs_Low_All = c(
    Low_Left = -0.5, Medium_Left = 0, High_Left = 0.5,
    Low_Right = -0.5, Medium_Right = 0, High_Right = 0.5
  ),

  # C2: Left vs Right (averaging across Low/Med/High)
  Left_vs_Right_All = c(
    Low_Left = 1/3,  Medium_Left = 1/3,  High_Left = 1/3,
    Low_Right = -1/3, Medium_Right = -1/3, High_Right = -1/3
  )
)
```

#### Contrast Definition

High Spin vs Low Spin

$$
C_1 = \frac{1}{2}(\mu_{\text{High\_Left}} + \mu_{\text{High\_Right}}) 
    - \frac{1}{2}(\mu_{\text{Low\_Left}} + \mu_{\text{Low\_Right}})
$$

Left vs Right Handed batters

$$
C_2 = \frac{1}{3}(\mu_{\text{Low\_Left}} + \mu_{\text{Medium\_Left}} + \mu_{\text{High\_Left}})
    - \frac{1}{3}(\mu_{\text{Low\_Right}} + \mu_{\text{Medium\_Right}} + \mu_{\text{High\_Right}})
$$

#### Unadjusted Comparison

```{r Contrast Definition}
summary(multcomp::glht(lm1, linfct = mcp(Combo = K_combo)), test = adjusted("none"))
```

Statistical Significance: (3 contrasts had significance) -
High_vs_Others_Laeft (p = .023) - High_vs_Low_All (p = .040) -
Left_vs_Right_All (p= .041)

High Spin means slightly lower Exit Velocity (.5-.8 mph) Right sided
batters have slightly higher Exit Velocity (.4 mph)

#### Bonferroni Adjustment

```{r Bonferroni Comparison}
summary(multcomp::glht(lm1, linfct = mcp(Combo = K_combo)), test = adjusted("bonferroni"))
```

Bonferroni Correction - controls the rate of false positives when
performing multiple hypothesis tests on the same dataset

After applying the Bonferroni Correction for multiple comparisons, we
see that none of the contrasts measured are statistically significant.

Therefore we fail to reject the null hypotheses that any of the
contrasts effect Exit Velocity after accounting for multiple testing in
the Bonferroni Correction

#### Holm Adjustment

```{r Holm Adjustment}
summary(multcomp::glht(lm1, linfct = mcp(Combo = K_combo)), test = adjusted("holm"))
```

Holm Correction - Controls family-wise error rate while being less
conservative and thus allowing for more statistical rejections.

After applying the Holm Correction for multiple comparisons, we see that
none of the contrasts measured are statistically significant.

Therefore we fail to reject the null hypotheses that any of the
contrasts effect Exit Velocity once the family-wise error was controlled
in the Holm Correction

#### Tukey's HSD

```{r Tukeys HSD}
fit_combo <- aov(ExitSpeed ~ Combo, data = df)
TukeyHSD(fit_combo)
```

Tukey's Honestly Significant Difference Test - is a Post-HOC test that
creates confidence intervals for all pairwise comparisons & controls for
the type 1 error rate to make sure that the differences are truly
significant.

The Tukey HSD test shows us that one pairwise comparison is
statistically significant (Low_Right & High_Left) which had a p value of
.022 and Low_Right having a mean difference of 1.10 mph more than
High_Left.

#### Synthesis/Discussion

While a few contrasts reached nominal significance before adjustment,
the magnitude of all observed effects was very small (\< 1 mph). Given
the large sample size (\~20,000 observations) and high within-group
variability (\~13 mph SD), these differences are statistically
detectable but not practically meaningful. Taken together, the contrast
and post-hoc results suggest that Combos have no substantial influence
on Exit velocity, and any minor variations are likely due to random
variation rather than true performance differences.

## Visualization

#### Boxplot/Stripchart

```{r Boxplot}
df$Combo <- factor(
  df$Combo,
  levels = c("Low_Left","Medium_Left","High_Left","Low_Right","Medium_Right","High_Right")
)

ggplot(df, aes(Combo, ExitSpeed, fill = Combo)) +
  geom_boxplot(alpha = 0.7, width = 0.7, outlier.shape = NA) +
  # (optional) stripchart look:
  geom_jitter(width = 0.15, alpha = 0.25, size = 1.3) +
  labs(title = "Exit Velocity by Combined Group (Spin Tier × Batter Side)",
       x = "Spin Tier × Batter Side", y = "Exit Velocity (mph)") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
  guides(fill = "none")
```

#### Mean +\_ SE Plot

```{r Mean SE Plot}
sumdat_combo <- df %>%
  group_by(Combo) %>%
  summarise(
    mean = mean(ExitSpeed),
    se   = sd(ExitSpeed)/sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(
    Combo = factor(Combo, levels = c("Low_Left","Medium_Left","High_Left",
                                     "Low_Right","Medium_Right","High_Right"))
  )

ggplot(sumdat_combo, aes(Combo, mean)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.15) +
  labs(title = "Mean ± SE of Exit Velocity (6 Combined Groups)",
       x = "Spin Tier × Batter Side", y = "Mean Exit Velocity (mph)") +
  theme_minimal(base_size = 13) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```
